---
title: "Software Review History"
format: 
  html:
    fig-width: 8
    fig-height: 4
    code-fold: false
---

```{r load-pkg, echo = FALSE}
library (dashboard)
library (ggplot2)
```


```{r get-rev-history, echo = FALSE}
#| cache: true
dat <- review_history (quiet = TRUE)
```

This chart shows the total number of packages under review each month, with
lines shown for all packages, as well as general review, and statistical
software review, which started in September 2021.


```{r history-months, echo = FALSE}
# Convert opened/closed dates to monthly sequence This is adapted from the editors
# timeline code in R/editors.R.
get_monthly_reviews <- function (dat) {
    dat$closed_at [which (is.na (dat$closed_at))] <- Sys.Date ()
    month_seqs <- lapply (seq_len (nrow (dat)), function (i) {
        s <- seq (dat$opened_at [i], dat$closed_at [i], by = "1 month")
        format (s, "%Y-%m")
    })
    month_seq_table <- table (unlist (month_seqs))
    data.frame (
        month = lubridate::ymd (paste0 (names (month_seq_table), "-01")),
        reviews = as.integer (month_seq_table)
    )
}

monthly_reviews <- get_monthly_reviews (dat)
monthly_reviews$type <- "all"
monthly_reviews$lty <- 1L
monthly_reviews_gen <- get_monthly_reviews (dat [which (!dat$stats), ])
monthly_reviews_gen$type <- "general"
monthly_reviews_gen$lty <- 2L
monthly_reviews_stats <- get_monthly_reviews (dat [which (dat$stats), ])
monthly_reviews_stats$type <- "stats"
monthly_reviews_stats$lty <- 1L

reviews <- rbind (monthly_reviews, monthly_reviews_gen, monthly_reviews_stats)
```

```{r plot-review-history, echo = FALSE}
ggplot (reviews, aes (x = month, y = reviews, colour = type)) +
    geom_line (linetype = reviews$lty) +
    xlab ("Year") +
    ylab ("Nr. active reviews / month") +
    theme_minimal () +
    theme (
        legend.position = "inside",
        legend.position.inside = c (0.1, 0.8)
    )
```

The next chart shows the average duration of the review process, plotted
against the months in which each review started. (The data have been slightly
smoothed to aid visual display.)

```{r rev-duration, echo = FALSE}
rev_dur <- data.frame (
    quarter = lubridate::quarter (dat$opened_at, type = "date_first"),
    duration = dat$duration_days * 12 / 365
) |>
    dplyr::group_by (quarter) |>
    dplyr::summarise (dur = mean (duration, na.rm = TRUE))
# Those quarterly data are quite jumpy, so apply a filter:
filter_dat <- function (dat, flen = 5) {
    index <- seq_len (length (dat) - flen + 1) + floor (flen / 2)
    x <- seq_len (flen) - ceiling (flen / 2)
    x <- -abs (x)
    filt <- exp (x)
    filt <- filt / sum (filt)
    dat [index] <- filter (dat, filt) [index]
    return (dat)
}
rev_dur$dur <- filter_dat (rev_dur$dur, flen = 3)
```

```{r rev-duration-plot, echo = FALSE}
ggplot (rev_dur, aes (x = quarter, y = dur)) +
    geom_line () +
    theme_minimal () +
    xlab ("Year") +
    ylab ("Review Duration (months)")
```
